---
title: "Bootstrapping"
author: "Melinda K. Higgins, PhD."
date: "February 25, 2017"
output: html_document
---

## Run Linear Regression - Extract adjusted R2 statistic

Given the formula and expected distribution of the adjusted R2, compute the 95% confidence interval for the adjusted R2 linear model fit statistic.

Not sure how to? This is not something you typically learn in an introductory statistics course.

That is ok, we can use bootstrapping to generate a lot of random linear models based on random samples using sampling with replacement. Then we generate a lot of estimates of the adjusted R2. We can then use these estimates of the adjusted R2 to compute the mean and a 95% confidence interval.

This is based on an example at the Quick-R website [http://www.statmethods.net/advstats/bootstrapping.html](http://www.statmethods.net/advstats/bootstrapping.html).

```{r setup}
knitr::opts_chunk$set(echo = TRUE)

# fit a linear model
m1 <- lm(mpg~wt+disp, data=mtcars)

# get the summary model fit statistics
sm1 <- summary(m1)

# extract adjusted R2 from model fit stats
sm1$adj.r.squared
```

## Setup Bootstrapping

```{r}
# Bootstrap 95% CI for adjusted R-Squared
library(boot)

# function to obtain R-Squared from the data 
adjrsq <- function(formula, data, indices) {
  d <- data[indices,] # allows boot to select sample 
  fit <- lm(formula, data=d)
  return(summary(fit)$adj.r.square)
} 

# bootstrapping with 1000 replications 
results <- boot(data=mtcars, statistic=adjrsq, 
  	R=1000, formula=mpg~wt+disp)
```

Of the 1000 boostrapped estimates, the average adjusted r2 is `r mean(results$t)`.

## First 20 bootstrapped adjusted R2 values

```{r}
# view results
knitr::kable(as.data.frame(results$t[1:20,]))
```

## Plot the distribution of the bootstrapped estimates

```{r}
# get bootstrapped results summary
results 

# plot the distribution of the boostrapped
# adjusted R2 values indicated by the 
# test statistic labeled "t"
plot(results)
```

## Get the 95% Confidence Intervals for the Bootstrapped Estimate of adjusted R2

```{r}
# get 95% confidence interval 
bci <- boot.ci(results, type="bca")
bci
```

