---
title: "HTML Widget - DT Example"
author: "Melinda K. Higgins, PhD."
date: "February 23, 2017"
output: html_document
---

## HTML Widgets

`htmlwidgets` [http://www.htmlwidgets.org/index.html](http://www.htmlwidgets.org/index.html) are helpful functions that allow you to embed interactive visualizations in HTML (Rmarkdown) reports.

To get ideas and see how these functions work and can be used, see the gallery at [http://www.htmlwidgets.org/showcase_leaflet.html](http://www.htmlwidgets.org/showcase_leaflet.html) and [http://gallery.htmlwidgets.org/](http://gallery.htmlwidgets.org/).

## Setup and Read-in Data

The code chunk below sets up all code chunks to be displayed in the final output (`echo = TRUE`) and loads the `tidyverse` package.

```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

Let's read-in the `Pesticides` dataset we used in the previous lecture. We'll then do the following:

1. compute the `AVG_ESTIMATE` from the `LOW_ESTIMATE` and `HIGH_ESTIMATE` provided in the original dataset _(the `mutate()` step)_;

2. `select()` the `COMPOUND, STATE_CODE, AVG_ESTIMATE` columns;

3. `group_by()` group the dataset by the `COMPOUND`s - we're using this to prepare statistical summaries of the amounts of these 388 pesticides used in the US during 2014;

4. finally we `summarise()` some of the descriptive statistics of the `COMPOUND`s - specifically we are capturing the median and 1st and 3rd quartiles (i.e., the limits of the interquartile range; 25th and 75th percentiles) - these are useful statisics since we know from our previous review of this dataset that many of the pesticide levels were heavily right-skewed - so reporting non-parametric statistics (median, Q1, Q3) is appropriate here.

```{r adddata}
# read in the pesticide dataset
Pesticides <- read_csv("Pesticides.csv")

pest1 <- Pesticides %>% 
  mutate(AVG_ESTIMATE = (LOW_ESTIMATE+HIGH_ESTIMATE)/2) %>%
  select(COMPOUND, STATE_CODE, AVG_ESTIMATE) %>%
  group_by(COMPOUND) %>%
  summarise(Q1 = quantile(AVG_ESTIMATE, 0.25, na.rm=TRUE),
            medianPest = median(AVG_ESTIMATE, na.rm=TRUE),
            Q3 = quantile(AVG_ESTIMATE, 0.75, na.rm=TRUE))

# see how big this summary table/data object is
dim(pest1)
```

Now that we have these summary stats for each of the Pesticides reported in this dataset, instead of using the `knitr::kable()` command to make a **STATIC** display table - let's instead used the `DT` package and the `datatable()` HTML widget to display **INTERACTIVELY** the information in this summary table.

### TABLE: Summary Statistics for Pesticides Detected in the US During 2014

For the interactive table shown below, feel free to:

* sort the columns by clicking on the arrows at the top of each column;
* make the table longer (display more rows) by clicking on the pull-down for "Show Entries"; or 
* search for a compound by name - begin by typing the beginning of the pesticide compound and the the table will begin to narrow down to selections matching what you're typing.

```{r DTdemo}
library(DT)
DT::datatable(pest1, options = list(pageLength = 5))
```

## Another fun HTML widget `radarchart`

Another fun HTML widget that is useful for visualizing multiple dimensions using a 2D graphic. See the demos for `radarchart` at [http://gallery.htmlwidgets.org/](http://gallery.htmlwidgets.org/) - scroll towards the bottom of the page. The Github repository for this widget is [https://github.com/mangothecat/radarchart](https://github.com/mangothecat/radarchart) which includes some demo code and examples in the README file.

Let's take the `Pesticides` dataset and look at the levels of 2 pesticides ("Glyphosate" and "Dichloropropene") in 6 counties in Florida in 2014. Let's make a **RADAR** plot of the levels of these 2 pesticides for the 6 counties where each county gets their own axis - so this **RADAR** chart has 6 axes. The plot is also **INTERACTIVE** so that when you "mouse-over" the points, a small pop-up box is shown with the levels of the 2 compounds at that point.

```{r}
FL_pest <- Pesticides %>%
  mutate(AVG_ESTIMATE = (LOW_ESTIMATE+HIGH_ESTIMATE)/2) %>%
  filter(STATE_CODE == 12) %>%
  filter(COUNTY_CODE %in% c(9,13,15,33,35,43)) %>%
  filter(COMPOUND %in% c("Glyphosate",
                         "Dichloropropene")) %>%
  select(COMPOUND, 
         COUNTY_CODE, 
         AVG_ESTIMATE)

FL_pest_wide <- FL_pest %>%
  spread(key = COMPOUND,
         value = AVG_ESTIMATE)

# link in the county names for Florida
library(choroplethrMaps)
data("county.regions")
countyFL <- county.regions %>%
  filter(state.fips.character == "12") %>%
  mutate(COUNTY_CODE = region - 12000)

FL_pest_wide2 <- left_join(FL_pest_wide, 
                           countyFL,
                           by="COUNTY_CODE")

FL_pest_wide2 <- FL_pest_wide2[,c(18,2:15)]

library(radarchart)
FL_pest_wide2 <- FL_pest_wide2 %>%
  select(county.name,
         Dichloropropene,
         Glyphosate)

radarchart::chartJSRadar(FL_pest_wide2, 
                         maxScale = max(FL_pest_wide2[,-1], 
                                        na.rm=TRUE), 
                         showToolTipLabel=TRUE)
```

```{r, echo=FALSE, eval=FALSE}
# these code chunks hide the code and it is not run
scores <- data.frame("Label"=c("Communicator", 
                               "Data Wangler", 
                               "Programmer", 
                               "Technologist",  
                               "Modeller", 
                               "Visualizer"),
                     "Rich" = c(9, 7, 4, 5, 3, 7),
                     "Andy" = c(7, 6, 6, 2, 6, 9),
                     "Aimee" = c(6, 5, 8, 4, 7, 6))

radarchart::chartJSRadar(scores, 
                         maxScale = 10, 
                         showToolTipLabel=TRUE)
```

